### 2023-02-05 article3.qmd
update

@@

```{r, load libraries}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
library(reactablefmtr)
library(gt)
```

@@

@@

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps. Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase. Then, we computed the z-score (i.e. standard score) of each slope, that follows a normal distribution. The detection of jumps thus consisted in spotting outliers in the z-score distribution. For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles. Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$. We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.

@@

# add levels
cflux$park_depth <- factor(cflux$park_depth, levels = c('200 m', '500 m', '1000 m'))

@@

# add level for order
#uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))

@@

At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to October with a sharp decrease in November. Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (< 0.5 particle/L) on the entire timeserie.

@@

@@

@@

# order floats by couple of deployment
particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))

@@

@@

# reload + redo some annoying stuff that were lost during the writing of the csv
dd <- vroom('heatmap_data.csv') |> mutate(Y = as.character(Y)) 
dd$X <- factor(dd$X, levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))
dd$Y <- factor(dd$Y, levels = c('102', '128', '161', '203', '256', '323', '406', '512', '645', '813', '1020', '1290'))
dd$Y <- factor(dd$Y, levels = c('1290', '1020', '813', '645', '512', '406', '323', '256', '203', '161', '128', '102'))
dd$wmo <- factor(dd$wmo, levels = c(6904240,4903634,6904241,1902578))

@@

### Combining OST and UVP data

The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e. 12.5 and 3.81). @Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region. In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt]. @eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj]. Due to those reasons, we have chosen to firstly compare $F_{small}$ and $F_{big}$ with UVP6 particles concentrations.

>>>>

+ dire que le but du papier est de faire une comparaison OST UVP qui sont deux instruments qui permettent de mesurer la meme chose (le flux des particules, l'abondance de particules) et donc le flux de carbone de manière indirecte
papier GRL Lacour de 2015

@@

```{r, load libraries}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
library(reactablefmtr)
library(gt)
library(gtExtras)
```

@@

all_floats <- all_floats |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))

@@

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps. Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase and computed the z-score (i.e. standard score) of each of them. The detection of jumps therefore consisted in spotting outliers in the z-score distribution. For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles. Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$. We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.

@@

# # add levels
# cflux$park_depth <- factor(cflux$park_depth, levels = c('200 m', '500 m', '1000 m'))

@@

# add level for order
#uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
# uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))

@@

At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to September with a sharp decrease starting in October. Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (< 0.5 particle/L) on the entire timeserie.

@@

TODO : It should be noted that all floats measure the same trend -> logique si on parle de la dynamique du bruit de fond de la zone et pas de la dynamique évènementielle qui serait différentes en fonction des flotteurs vu qu'ils dérivent chacun dans le parcours respectifs. Ça devrait renforcer cet aspect.

### Comparing OST and UVP data

The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e. 12.5 and 3.81). @Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region. In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt]. @eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj]. As a result, the direct comparison of carbon fluxes derived from both OST and UVP data with @eq-guidi-article3 should be focused on the trend rather than on absolute carbon flux values.

```{r, compute guidi fluxes at parking depth}
#| label: fig-comparison-ost-uvp
#| fig-cap: Comparison of total carbon fluxes derived from the OST (black) and the UVP6 (red) where the error bars represent the 25$^{th}$ and 75^{th}$ percentiles around the median.

compute_flux <- function(A, b, PSD){ # PSD = Particle Size Distribution
  # for classes sizes between 0.102 mm and 1.5 mm
  #mid_ESD <- c(0.1164512, 0.1463600, 0.1843915, 0.2325200, 0.2933257, 0.3691650, 0.4650400, 0.5860455, 0.7385533, 0.9280422, 1.1705684, 1.4795321) # With Rainer's code (paruvpy)
  # mid_ESD <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
  #              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
  #              1.16170742, 1.46365963) # With Lionel's script
  mid_ESD <- c(0.23051195, 0.29042685,
               0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
               1.16170742, 1.46365963) # With Lionel's script
  exp <- A * mid_ESD ** b
  estimated_flux <- rowSums(t(t(PSD)*exp))
  return(estimated_flux)
}

parking_data <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
parking_data <- parking_data |> filter(PhaseName == 'PAR') |> drop_na(NP_Size_102) |> select(park_depth, all_of(lpm_class), dates = juld, park_depth, wmo, cycle) |> mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578)),
                                   park_depth = factor(park_depth, levels = c('200 m', '500 m', '1000 m')))
parking_spectra <- parking_data |> dplyr::select(NP_Size_203:NP_Size_1290)

parking_data <- parking_data |> mutate(guidi_flux = compute_flux(12.5, 3.81, parking_spectra)) |> group_by(park_depth, cycle, wmo) #|> summarize(median_guidi_flux = median(guidi_flux))

# plot it
comparison_flux_ost_uvp <- merge(parking_data, cflux) |> mutate(total_flux = small_flux + big_flux)

# https://stackoverflow.com/questions/41077199/how-to-use-r-ggplot-stat-summary-to-plot-median-and-quartiles
# ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = small_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
#   geom_errorbar(aes(x = date, y = guidi_flux),
#                       stat = 'summary',
#                       fun.min = function(z) {quantile(z, 0.25)},
#                       fun.max = function(z) {quantile(z, 0.75)},
#                       fun = median, colour = 'red', width = 5) +
#   geom_line(aes(x = date, y = guidi_flux),
#                       stat = 'summary',
#                       fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') +
#   geom_line(data = comparison_flux_ost_uvp, aes(x = date, y = big_flux), linetype = 'dotdash') + theme_bw()

ggplot(comparison_flux_ost_uvp) + geom_line(aes(x = date, y = total_flux)) + facet_wrap(~wmo~park_depth, scales = 'free_y', nrow = 4) +
  geom_errorbar(aes(x = date, y = guidi_flux),
                      stat = 'summary',
                      fun.min = function(z) {quantile(z, 0.25)},
                      fun.max = function(z) {quantile(z, 0.75)},
                      fun = median, colour = 'red', width = 5) +
  geom_line(aes(x = date, y = guidi_flux),
                      stat = 'summary',
                      fun = median, colour = 'red') + scale_y_continuous(trans = 'log10') + theme_bw()

correlations <- comparison_flux_ost_uvp |> select(park_depth, wmo, cycle, guidi_flux, small_flux, big_flux, date, total_flux) |> group_by(wmo, park_depth, cycle) |> mutate(median_guidi_flux = median(guidi_flux)) |> group_by(park_depth, wmo) |> summarise(cor_small = round(cor(median_guidi_flux, small_flux),2), cor_big = round(cor(median_guidi_flux, big_flux),2), cor_total = round(cor(median_guidi_flux, total_flux),2))

correlations <- correlations |> pivot_wider(names_from = park_depth, values_from = c(cor_small, cor_big, cor_total))
gt(correlations) |> tab_spanner(label = "Small flux", columns = 2:4) |> tab_spanner(label = "Big flux", columns = 5:7) |> tab_spanner(label = 'Total flux', columns = 8:10) |>
  cols_label(wmo = 'WMO', `cor_small_200 m` = '200 m',
             `cor_small_500 m` = '500 m',
             `cor_small_1000 m` = '1000 m',
             `cor_big_200 m` = '200 m',
             `cor_big_500 m` = '500 m',
             `cor_big_1000 m` = '1000 m',
             `cor_total_200 m` = '200 m',
             `cor_total_500 m` = '500 m',
             `cor_total_1000 m` = '1000 m')

TODO :

- TS diagram at drifting depth + isopycnals
- NPQ chla + MLD (same y axis) 
- satellite chla + MLD -> show that conditions for bloom (stable and shallow MLD) + chla happened before deployment in the area
- faire le modèle de pente avec les biovolues pour prédire quels types de particules explique le plus la pente -> faire un modèle pour le small flux et un modèle pour le big flux, pas un modèle pour les small + big car les pentes ne sont pas additives
- discuter des corrélations
- science: on voit le système latent, on a raté le bloom
- rajouter les vitesses de sédimentations (n'importe quelle méthode fera l'affaire)
- rajouter dans les méthodes les produits satellites qu'on utilise
- différentes dynamiques post-bloom
- phénologie du bloom printanier
- MLD au cours du temps dans la zone, voir les variations

### Which particles drive

```{r, flux model 2, include = FALSE}

#load(file = 'DATA/MODEL/model_xgboost_data')

@@

Cette partie devrait venir dans la discussion si on part du principe que la première partie est liée à la méthodologie

@@

# order floats by couple of deployment
# particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))

@@

### Environmental conditions

Add NPQ plot avec MLD dessus, les memes échelles verticales (depth) + aller chercher les produits satellitaires pour montrer que l'on a bien raté le bloom ou que les conditions du bloom n'étaient pas réunies

@@

# reload + redo some annoying stuff that were lost during the writing of the csv
dd <- vroom('heatmap_data.csv') |> mutate(Y = as.character(Y)) 
dd$X <- factor(dd$X, levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))
dd$Y <- factor(dd$Y, levels = c('102', '128', '161', '203', '256', '323', '406', '512', '645', '813', '1020', '1290'))
dd$Y <- factor(dd$Y, levels = c('1290', '1020', '813', '645', '512', '406', '323', '256', '203', '161', '128', '102'))
# dd$wmo <- factor(dd$wmo, levels = c(6904240,4903634,6904241,1902578))

@@
